---
title: "My Portfolio for Computational Musicology 2025"
author: "Nora Kahleogullari"
date: "Block 4"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme:
      bg: "#e9fce6"
      fg: "#4a4949"
      primary: "#87f5a2"
      navbar-bg: "#86f0c4"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(flexdashboard)
library(plotly)
library(ggcorrplot)
library(tibble)
library(dbscan)
library(kableExtra)
library(randomForest)
library(caret)
#library(shiny)
source("compmus.R")
knitr::opts_chunk$set(echo = TRUE)
aisc2024 <- read_csv('compmus2025.csv')
```


Introduction
=================================================

The first song I created is a dark, 80's inspired R&B track I wanted to make something similar to the Weeknd's music, because I always loved his ability to blend nostalgia with a modern edge. I tried to implement synths and R&B to create his atmospheric yet energetic sound. My goals was to capture the listeners emotionally, while still having a strong rhythm and melody. 

My prompt for the first song was and it is made with Udio.ai: 
song has a dark, atmospheric 80s-inspired synthwave beat with smooth R&B elements, like The Weeknd , moody, energetic, downtempo

'<audio controls>
  <source src="nora-k-1.mp3" type="audio/mpeg">
  Your browser does not support the audio element.
</audio>'

For the second song I decided to create a Latin summer song because I like the vibrant energy and the strong rhythms. There's something about Latin music and instantly puts you in a good mood, it's lively and perfect for warm summer days. I wanted to capture that feeling by incorporating upbeat percussion and a rhythmic guitar. My goal was to create a track that people can enjoy most of the time in summer. 

My second song was made with Suno.ai and my prompt for the song was:
The second song is dembow, reggaeton, cumbia, afro-latin fusion, rap, latin trap. It's like summer song that will bring you the back to the summer

'<audio controls>
  <source src="nora-k-2.mp3" type="audio/mpeg">
  Your browser does not support the audio element.
</audio>'


The table below presents the extracted Essentia feature values for my generated tracks.
```{r echo=FALSE}
mycorpus <- aisc2024 |>
  filter(filename %in% c('nora-k-1','nora-k-2'))
  
kable(mycorpus, caption = "Essentia Feature Analysis of My Tracks") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```


Discovery
================================================

Column {.sidebar data-width=400}
-----------------------------------------------
### description 
First we have a few plots, so a better insight into the data can be created


The histogram of danceability shows that most songs are moderately danceable


The plots show us that a higher tempo does not indicate an higher danceability. While it may seem suprising, danceability depends on multiple factors, including, rhythm, valence, instrumentation and arousal, rather than just tempo. 

A high-tempo song (e.g., EDM) with poor rhythm may feel less danceable.. The definition of dancability also matters - if it strictly means jumping to the bet, the interpretation would change. A song with high tempo but low energy may not feel as danceable as a moderate-tempo track with a strong rhythm and energy.

Tempo distribution reveals a tendency toward either lower BPM values or BPM's above 110. This aligns with expectations: dance, music typically exceeds 110 BPM, while ballads and classical pieces tend to be slower.

The boxplot suggests a positive correlation between danceability and valence, whereas instrumentalness appears independant of danceability. 
The dataset contains many high-danceability, high-valence tracks.

The interactive chart allows you to hover for additional details. It shows that arousal hovers around 4, and while danceability fluctates, arousal remains fairly stable. This chart resembles the boxplot but focuses on arousal instead of valence. 

The heatmap reveals the following correlation:
- A positive relationship between engagement and arousal, dancebility and arousal and engagement and dancebility
- A negative correlation between instrumentalness and arousal

There is no clear correlation between arousal en engaginess and whether the songs are AI-generated or not


Column {data-width = 1100}
----------------------------------------------------

```{r echo=FALSE}
aisc2024 <- read_csv('compmus2025.csv')

# Histogram of Danceability
ggplot(aisc2024, aes(x = danceability)) +
  geom_histogram(binwidth = 0.05, fill = "purple", alpha = 0.7, color = "black") +
  labs(title = "Histogram of Danceability", x = "Danceability", y = "Frequency") +
  theme_minimal()


# Scatter Plot of Danceability vs Tempo
ggplot(aisc2024, aes(x = tempo, y = danceability)) +
  geom_point(alpha = 0.7, color = "purple") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Danceability vs. Tempo", x = "Tempo (BPM)", y = "Danceability") +
  theme_minimal()
```


```{r pressure, echo=FALSE}

ggplot(aisc2024, aes(x=tempo)) +
  geom_histogram(binwidth=10, fill="steelblue", color="black", alpha=0.7) +
  theme_minimal() +
  labs(title="Distribution of Tempo",
       x="Tempo (BPM)",
       y="Count") +
  geom_density(aes(y=..count..*10), color="red", lwd=1.2)

aisc2024$valence <- cut(aisc2024$valence, 
                          breaks = c(0, 0.33, 0.66, 1), 
                          labels = c("Low", "Medium", "High"))

# Box plot: Danceability across Valence levels, colored by Instrumentalness
ggplot(aisc2024, aes(x = valence, y = danceability, fill = valence)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(aes(color = instrumentalness), width = 0.2, alpha = 0.6) +  
  scale_fill_manual(values = c("Low" = "blue", "Medium" = "orange", "High" = "red")) +
  scale_color_gradient(low = "black", high = "white") +  
  theme_minimal() +
  labs(title = "Danceability Across Valence Levels (Instrumentalness in Points)",
       x = "Valence Category",
       y = "Danceability",
       color = "Instrumentalness") +
  theme(legend.position = "right")
```


```{r echo=FALSE}
library(dplyr)
library(ggplot2)
library(plotly)

p <- aisc2024 |>
  ggplot(
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point(alpha = 0.7) +  # Regular points
  
  # Highlight Track 1
  geom_point(
    data = aisc2024 |> filter(tempo == 111, arousal == 5.0956),  
    aes(size = 7),  
    colour = "red",
    shape = 21,
    stroke = 1.5
  ) +
  
  # Highlight Track 2
  geom_point(
    data = aisc2024 |> filter(tempo == 134, arousal == 5.34),  
    aes(size = 7),  
    colour = "blue",
    shape = 21,
    stroke = 1.5
  ) +
  
  geom_rug(linewidth = 0.2, sides = "b", alpha = 0.5) +  
  geom_text(
    x = 111,
    y = 5.0956,
    label = "Track 1",
    size = 4,
    hjust = "left",
    vjust = "center",
    angle = 25,
    colour = "black"
  ) +
  geom_text(
    x = 134,
    y = 5.34,
    label = "Track 2",
    size = 4,
    hjust = "left",
    vjust = "center",
    angle = 25,
    colour = "black"
  ) +
  # scale_x_continuous(
  #   limits = c(50, 200),
  #   breaks = seq(50, 200, 50),
  #   minor_breaks = NULL
  # ) +
  scale_y_continuous(
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c(option = "plasma", direction = -1) +
  scale_size_continuous(trans = "sqrt", guide = "none") +
  theme_minimal(base_size = 14) +
  labs(
    title = "How Tempo and Arousal Relate in the Playlist",
    subtitle = "Higher tempo generally correlates with higher arousal, with danceability as color",
    x = "Tempo (BPM)",
    y = "Arousal (1-9 Scale)",
    colour = "Danceability"
  )

ggplotly(p)

```

```{r echo=FALSE}
corr_matrix <- cor(aisc2024[, -c(1,8)]) 
ggcorrplot(corr_matrix, method = "circle")
```

```{r echo=FALSE, fig.height=5, fig.weigth=7}
ggplot(aisc2024, aes(x = engagingness, y = arousal)) +
  geom_point() +
  facet_wrap(~ ai) +  
  theme_minimal()
```


Insights to the whole corpus
=================================================
Column {.sidebar data-width=350}
-----------------------------------------------------
### Description
The songs in the corpus generally maintain a moderate tempo, as evidenced by a broad peak around 100 BPM

In the first song, there are two noticeable tempo drops around 50 and 90 seconds, with additional tempo variations in the outro.

The second song maintains a steady tempo until 110 seconds, and after it stays steady until 145 seconds in the outro 


Column {data-width=1150}
------------------------------------------------
Tempi
```{r echo=FALSE}
df <- read_csv("compmus2025.csv")

ggplot(df, aes(x = tempo)) +
  geom_histogram(bins = 10, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Tempi for Class Corpus",
       x = "Tempo (BPM)",
       y = "Count") +
  theme_minimal()
```


My contributions 

```{r echo=FALSE, fig.height=6, fig.width=4}

"features/nora-k-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

"features/nora-k-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
```



Key profile
================================================

Column {.sidebar data-width = 400}
------------------------------------------------------

### Description
song 1 chord
The frequent use of the seventh chord indicates a blues influence, as it's used to add harmonic richness and tension. The repetition of chord sequences indicates the theme that is present in the whole song. 

Song 2 chord 
As can be seen there are many chords used, as you can hear. The structure can be seen throughout the song. The intro is from 0-25s then the verse 25-48s, chorus 48-75s verse 2 75-98s, chorus again 98-125s and then from 175s the outro.

Column {data-width=1100}
-------------------------------------------
### Song 1
```{r echo=FALSE}
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(5.0, 2.0, 3.5, 2.5, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 2.0, 2.5)

minor_key <-
  c(5.0, 2.5, 3.5, 4.5, 2.0, 3.5, 2.0, 4.0, 3.5, 2.5, 3.0, 2.5)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

"features/nora-k-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "identity",       # Try different norms (and match it with what you used in compmus_chroma)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()   

```



### song 2
```{r echo=FALSE}
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(5.0, 2.0, 3.5, 2.5, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 2.0, 2.5)

minor_key <-
  c(5.0, 2.5, 3.5, 4.5, 2.0, 3.5, 2.0, 4.0, 3.5, 2.5, 3.0, 2.5)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

"features/nora-k-2.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "identity",       # Try different norms (and match it with what you used in compmus_chroma)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()   
```



Deep dive into the music
============================================
Column {.sidebar data-width=400}
--------------------------------------------------

### Description
track 1
chroma: The R&B song uses overall only 2 pitchclasses, while in the middle part it almost uses every chord.

self-similarity:
It can be seen that the R&B track has a lot more similar parts in the song, as there are a lot of little black squares. This is because the clap and the synths are coming back with a repeated interval.

timbre-based similarity: The fist part of the song is, in terms of similarity, super similar until the chorus, and after its again very similar but there are also a bit more changes.


track 2
chroma :
This song shows a sustained activity across all pitch classes. It shows that this song is more intricate and has a complexer harmonic progression

self-similarity: The latin track seems different but also the same, because it seems that the green lines are almost woven, this is because the beat is very similar but it has a sort of distortion above it. 

timbre-based similarity: In the intro of this song, the song doesn't change much, but it does differ a lot from the rest of the song, the rest of the song has big chunks that are also similar except for the outro and a small part around 110 seconds.


When comparing the timbre-based self similarity matrix and the tempogram we can see that they show on similar places the difference.

Column {.tabset}
------------------------------------------------------
### Track 1 


####### chromagram
```{r echo=FALSE, fig.height=5, fig.width=5}
"features/nora-k-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL, title="Chromagram track 1") +
  theme_classic()                                     

```


#######  self-similarity
```{r, echo=FALSE, fig.height=5, fig.width=5}
"features/nora-k-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL,
       title="self-similarity track 1") +
  theme_classic()
```


####### timbre-based similarity
```{r, echo=FALSE, fig.height=5, fig.width=5}
"features/nora-k-1.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL,
       title="timbre based similarity track 1") +
  theme_classic()   
```


### Track 2

####### Chromagram
```{r, echo=FALSE, fig.height=5, fig.width=5}
#### Chroma - Track 2

"features/nora-k-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL,
       title="Chromagram track 2") +
  theme_classic()
```


#######  self-similarity
```{r echo=FALSE, fig.height=5, fig.width=5}
"features/nora-k-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL,
       title="self similarity track 2") +
  theme_classic()
```


#######  timbre based self similarity
```{r echo=FALSE, fig.height=5, fig.width=5}
"features/nora-k-2.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL,
       title="timbre based similarity track 2") +
  theme_classic()   

```


Power Novelty
===============================================
Column {.sidebar data-width=400}
------------------------------------------------
### Description 

Track 1:
The first peak in the energy novelty is at the beginning just before the lyrics starts, when you hear a drum/buzz that indicates when the real beats starts. The second peak is around 55seconds when the beat changes and it becomes lighter and less heavy. The third peak is also seen in the spectral novelty when the beats begins to start again.


Track 2:
In the energy novelty graph, there are 3 high peaks, indicating percussive detection, the first is when the music comes in, the second peak is again around 55secons its when she says a bacho, so that is a harsh sound. The final peak see we also in the spectral novelty graph. It shows that around 110-120 seconds there is a drop, thats because there is an instrumental break before the lyrics an beat start again.

Column {.tabset}
------------------------------------------------

### Track 1

```{r, echo=FALSE, fig.height=5, fig.width=4}
"features/nora-k-1.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty", title="Energy Novelty Track 1")



"features/nora-k-1.json" |> 
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty", title="Spectral Novelty Track 1")
```



### Track 2

```{r echo=FALSE, fig.height=5, fig.width=4}
"features/nora-k-2.json" |>
  compmus_energy_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty", title="Energy Novelty Track 2")



"features/nora-k-2.json" |> 
  compmus_spectral_novelty() |> 
  ggplot(aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Spectral Novelty", title="Spectral Novelty Track 2")
```



Visualizing the predicting power
================================================

Column {.sidebar data-width=400}
-----------------------------------------------
### Description 

When optimizing the RandomForest it improved. It improved with the precision and recall of AI. For non-AI it slightly differs each time but the difference is not very big


From the dataframe it can be seen that danceability and approachability are far out the best features to potentially generate the most meaningfull clusters. 

Column {data-width = 1100}
----------------------------------------------------

```{r echo=FALSE}
compmus2025 <- read_csv("compmus2025.csv")

get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
} 

cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))

classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) 

compmus_cv <- compmus2025_filtered |> vfold_cv(5)

forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")
indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit_resamples(
    compmus_cv, 
    control = control_resamples(save_pred = TRUE)
  )


indie_forest |> get_conf_mat() |> autoplot(type = "heatmap")
```


```{r}
indie_forest |> get_pr()

```

```{r}

forest_model <-
  rand_forest() |>
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")
indie_forest <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model)

indie_forest_resampled <- indie_forest |>
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))


indie_forest_resampled |> get_conf_mat() |> autoplot(type = "heatmap")
```

```{r}
indie_forest_resampled |> get_pr()

```

```{r}
workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(forest_model) |> 
  fit(compmus2025_filtered) |> 
  pluck("fit", "fit", "fit") |>
  ranger::importance() |> 
  enframe() |> 
  mutate(name = fct_reorder(name, value)) |> 
  ggplot(aes(name, value)) + 
  geom_col() + 
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Importance")
```


Clustering
================================================

Column {.sidebar data-width=400}
-----------------------------------------------
### Description 


Clustering using the single linkage method yields the least favorable result, with very small cluster splits and an imbalance in vertical alignment. The average linkage method improves upon this, as we can observe larger clusters higher up in the dendrogram. However, on the left side, there remains a cluster that is not accounted for in the higher-level clusters. The complete linkage method produces the most balanced clusters, where we can see that the first two clusters at the top contain all the songs.

From the tibble, it is evident that when the data is divided into 8 clusters, the distribution is mostly even across the clusters.



Column {data-width = 1100}
-------------------------------------------------

```{r echo=FALSE}
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  

aisc2024$valence <- as.numeric(as.character(aisc2024$valence))

cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = aisc2024
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(aisc2024) |>
  juice() |>
  column_to_rownames("filename")


compmus_dist <- dist(cluster_juice, method = "euclidean")

compmus_dist |> 
  hclust(method = "average") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()
```



```{r echo=FALSE}
library(cluster)

silhouette_score <- function(dist_matrix, clustering) {
  silhouette(clustering, dist_matrix) |> summary()
}

compmus_hclust <- hclust(compmus_dist, method = "average") # Adjust method

cluster_assignments <- cutree(compmus_hclust, k = 8) # Change k as needed
silhouette_score(compmus_dist, cluster_assignments)
```




Exclusive insights
=========================================

Both songs exhibit a highly structured composition, likely due to the fact that they were generated by artificial intelligence. Since AI models often rely on predefined patterns or algorithms to create music, it is reasonable to assume that a formulaic approach was used to construct these pieces. Additionally, a clear pattern emerges when analyzing the spectral novelty and energy novelty of both songs. Notably, both pieces display peaks in the exact same locations, reinforcing the idea that they were created using a systematic method. This consistency suggests that the AI followed a structured framework when generating the music, ensuring a balanced and cohesive flow throughout each song.