---
title: "My Portfolio for Computational Musicology 2025"
author: "Nora Kahleogullari"
date: "Block 4"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(plotly)
library(tibble)
#library(shiny)
source("compmus.R")
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to my dashboard!

This dashboard includes insights into music analysis, visualizations, and key musical features.

### Tab 1: Tempi and Music Features

Column {data-width=300}
-----------------------------------------------------------------------------------------

### Insights

The first song has a dark, atmospheric 80s-inspired synthwave beat with smooth R&B elements, like The Weeknd, moody, energetic, downtempo, made with Udio.ai.

The second song is dembow, reggaeton, cumbia, afro-latin fusion, rap, latin trap. It's like a summer song that will bring you back to the summer, made with Suno.ai.

#### Song Analysis

```{r echo=FALSE}
"features/nora-k-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme? 
```

### Tab 2: Discovery and Insights into Arousal

Column {data-width=300}
------------------------------------------------------------------------------------------

### More Insights

The dataset contains several features related to danceability, tempo, and arousal. This section focuses on the relationships between these features.

#### Danceability and Tempo

```{r echo=FALSE}
aisc2024 <- read_csv('compmus2025.csv')

# Scatter Plot of Danceability vs Tempo
ggplot(aisc2024, aes(x = tempo, y = danceability)) +
  geom_point(alpha = 0.7, color = "purple") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Danceability vs. Tempo", x = "Tempo (BPM)", y = "Danceability") +
  theme_minimal()
```

### Analysis of the Data

Column {data-width=300}
------------------------------------------------------------------------------------------

#### Insights into Tempo

The histogram shows that most songs are moderately danceable. Higher tempo correlates with higher danceability, but not in every case. Songs that are more instrumental tend to have lower danceability.

---

### Interactive Plots

Column {data-width=300}
------------------------------------------------------------------------------------------

#### Chart 1

```{r echo=FALSE}
# Scatter Plot for Tempo vs Arousal, with Instrumentalness and Danceability
p <- aisc2024 |>
  ggplot(
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point(alpha = 0.7) +  # Regular points
  
  # Highlight Track 1
  geom_point(
    data = aisc2024 |> filter(tempo == 111, arousal == 5.0956),  
    aes(size = 7),  
    colour = "red",
    shape = 21,
    stroke = 1.5
  ) +
  
  # Highlight Track 2
  geom_point(
    data = aisc2024 |> filter(tempo == 134, arousal == 5.34),  
    aes(size = 7),  
    colour = "blue",
    shape = 21,
    stroke = 1.5
  ) +
  
  geom_rug(linewidth = 0.2, sides = "b", alpha = 0.5) +  
  geom_text(
    x = 111,
    y = 5.0956,
    label = "Track 1",
    size = 4,
    hjust = "left",
    vjust = "center",
    angle = 25,
    colour = "black"
  ) +
  geom_text(
    x = 134,
    y = 5.34,
    label = "Track 2",
    size = 4,
    hjust = "left",
    vjust = "center",
    angle = 25,
    colour = "black"
  ) +
  scale_y_continuous(
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c(option = "plasma", direction = -1) +
  scale_size_continuous(trans = "sqrt", guide = "none") +
  theme_minimal(base_size = 14) +
  labs(
    title = "How Tempo and Arousal Relate in the Playlist",
    subtitle = "Higher tempo generally correlates with higher arousal, with danceability as color",
    x = "Tempo (BPM)",
    y = "Arousal (1-9 Scale)",
    colour = "Danceability"
  )

ggplotly(p)
```

### Final Insights

Each tab in the `flexdashboard` now has a specific sidebar that provides context or insight relevant to the content displayed within the tab. The sidebar content will change dynamically when you switch between tabs.

This approach leverages `flexdashboard`'s ability to place sidebars under each individual section (`Column`). By customizing the sidebar content in each column, we ensure that the sidebar updates based on the active tab.